{\bf Dynamic Scheduling in Halide}

### 2013-12-17 15869/paper.tex
motivation

\usepackage{fullpage}
\usepackage{manfnt}
\usepackage{wasysym}
\usepackage{listings}
\usepackage{graphicx}
\usepackage{url}
%\usepackage{ulem}
\usepackage{marvosym}
%\usepackage{skull}
\usepackage{proof}
\usepackage{array}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Motivation}
% TODO

>>>>

\usepackage{caption,subcaption}
\usepackage{fullpage}
\usepackage{color}
\usepackage{manfnt}
\usepackage{wasysym}
\usepackage{listings}
\usepackage{graphicx}
\usepackage{url}
%\usepackage{ulem}
\usepackage{marvosym}
%\usepackage{skull}
\usepackage{proof}
\usepackage{array}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Motivation}

\definecolor{grey}{RGB}{127,127,127}
\definecolor{darkcyan}{RGB}{0,127,127}
\definecolor{olivegreen}{RGB}{0,127,0}
\definecolor{violet}{RGB}{127,0,127}
\definecolor{brickred}{RGB}{127,0,0}
\definecolor{brown}{RGB}{127,63,0}
\definecolor{orange}{RGB}{192,96,0}
\newcommand\hilight[2]{\color{#1}#2\color{black}}

\begin{figure}[t]
	\begin{center}
	\begin{subfigure}[b]{0.45\textwidth}
		\begin{center}
		\begin{tabular}{l}
		{\em User-provided schedule:} \\
		\\
		\hilight{blue}{\texttt{blur\_x /* nothing special */;}} \\
		\hilight{olivegreen}{\texttt{blur\_y.store\_root().compute\_root();}} \\
		\\
		{\em Compiler-generated code:} \\
		\\
                \texttt{\hilight{olivegreen}{float~blur\_y[HEIGHT][WIDTH];}} \\
                \texttt{\hilight{olivegreen}{for~(row~=~0~to~HEIGHT)~\{}} \\
                \texttt{\hilight{olivegreen}{~~for~(col~=~0~to~WIDTH)~\{}} \\
                \texttt{\hilight{olivegreen}{~~~~blur\_y[row][col] = ...;}}\\
                \texttt{\hilight{olivegreen}{~~\}}} \\
                \texttt{\hilight{olivegreen}{\}}} \\
		\\
                \texttt{\hilight{blue}{float~blur\_x[HEIGHT][WIDTH];}} \\
                \texttt{\hilight{blue}{for~(row~=~0~to~HEIGHT)~\{}} \\
                \texttt{\hilight{blue}{~~for~(col~=~0~to~WIDTH)~\{}} \\
                \texttt{\hilight{blue}{~~~~blur\_x[row][col] = ...;}}\\
                \texttt{\hilight{blue}{~~\}}} \\
                \texttt{\hilight{blue}{\}}} \\
		\end{tabular}
		\end{center}
		\caption{\texttt{blur\_x} and \texttt{blur\_y} are scheduled separately. This schedule, while easy to understand, offers poor locality (values computed are not used until a full image traversal later) and poor parallelism (creating one thread per row or column of pixels both floods the system with too many threads and creates cache contention).}
		\label{fig:scheds-a}
	\end{subfigure}
	\qquad
	\begin{subfigure}[b]{0.5\textwidth}
		\begin{center}
		\begin{tabular}{l}
		{\em User-provided schedule:} \\
		\\
                \hilight{blue}{\texttt{blur\_x.tile(x, y, x2, y2, 8, 8).parallel(y);}}\\
                \hilight{olivegreen}{\texttt{blur\_y.store\_at(blur\_x, x).compute\_at(blur\_x, y2);}}\\
                \\
		{\em Compiler-generated code:} \\
		\\
                \texttt{\hilight{blue}{float~blur\_x[HEIGHT][WIDTH];}} \\
                \texttt{\hilight{blue}{PARALLEL for~(row~=~0~to~8)~\{}} \\
                \texttt{\hilight{blue}{~~for~(col~=~0~to~8)~\{}} \\
                \texttt{\hilight{olivegreen}{~~~~float~blur\_y[HEIGHT/8 + 2][WIDTH/8 + 2];}} \\
                \texttt{\hilight{blue}{~~~~for~(row2~=~0~to~WIDTH/8 + 2)~\{}} \\
                \texttt{\hilight{olivegreen}{~~~~~~for~(col2~=~0~to~WIDTH/8 + 2)~\{}} \\
                \texttt{\hilight{olivegreen}{~~~~~~~~blur\_y[row2][col2]~=~...;}} \\
                \texttt{\hilight{olivegreen}{~~~~~~\}}} \\
                \texttt{\hilight{blue}{~~~~~~for~(col2~=~0~to~WIDTH/8)~\{}} \\
                \texttt{\hilight{blue}{~~~~~~~~blur\_x[row*8~+~row2][col*8~+~col2]~=~...;}} \\
                \texttt{\hilight{blue}{~~~~~~\}}} \\
                \texttt{\hilight{blue}{~~~~\}}} \\
                \texttt{\hilight{blue}{~~\}}} \\
                \texttt{\hilight{blue}{\}}} \\
		\end{tabular}
		\end{center}
		\caption{\texttt{blur\_x} is tiled in 8-fold splits in each dimension, and \texttt{blur\_y} is scheduled at the tile level within \texttt{blur\_x}. This schedule offers good locality (values are used soon after being computed) and good parallelism (one thread per tile; tile size can be customized).}
		\label{fig:scheds-b}
	\end{subfigure}
	\end{center}
	\caption{Two example function schedules for the two-stage blur algorithm shown in Figure~\ref{fig:intro}.}
	\label{fig:scheds}
\end{figure}

The high concept of Halide is to separate the logic of an image processing algorithm from the order in which the functions for each stage are scheduled. Figure~\ref{fig:scheds} shows two possible schedules for our example algorithm, and discusses the performance characteristics of each.
Though Figure~\ref{fig:scheds-b} is more cache-friendly and offers superior paralellism opportunity, a notable disadvantage is that some computation is duplicated across tiles. This is shown in the extra 2 pixels computed for \texttt{blur\_y} in each dimension. In this example, the overhead is minor, as Halide can statically reason about the maximum extra ``padding'' that \texttt{blur\_y} must compute for each tile of \texttt{blur\_x}.

However, this extra computation can become a significant problem when the dependencies between the algorithm's stages are computed at runtime and hence are not available for compile-time reasoning, as in the PatchMatch algorithm~\cite{patchmatch} and in edge detection. If the user wishes to use a cache-friendly, highly-parallel function schedule, as in Figure~\ref{fig:scheds-b}, Halide must make a conservative estimate regarding how much extra padding is required, resulting in extra work duplication.

To get a sense for how much extra unnecessary computation can arise, we introduce a sample two-stage Halide program in which the second stage (\texttt{flip()}) dynamically computes which pixels it requires from the first stage (\texttt{invert()}). This program is described in Figure~\ref{fig:flip}. We then ran this program on a test image using the simple schedule from Figure~\ref{fig:scheds-a} and using several nested schedules as in Figure~\ref{fig:scheds-b}. We found that in the former situation, 13,068 calls to \texttt{invert()} were required, while in the latter, \texttt{invert()} was invoked anywhere between 252,648 to 7,847,400 times depending on the schedule used.
Hence, the motivation for {\em dynamic scheduling} is to extend Halide's ability to avoid duplicate work to apply to this wider range of algorithms.

\begin{figure}[t]
	\begin{tabular}{cc}
	\begin{subfigure}[b]{0.45\textwidth}
		\begin{center}
		\includegraphics[width=0.85\textwidth]{canoe-0.png}
		\end{center}
		\caption{Original input image.}
	\end{subfigure} &
	\begin{subfigure}[b]{0.45\textwidth}
		\begin{center}
		\includegraphics[width=0.85\textwidth]{canoe-1.png}
		\end{center}
		\caption{Stage 1 is a simple value-invert of each input pixel.}
		\label{fig:invert}
	\end{subfigure} \\
	& \\
	\begin{subfigure}[b]{0.45\textwidth}
		\begin{center}
		\includegraphics[width=0.85\textwidth]{canoe-2.png}
		\end{center}
		\caption{The core of stage 2 is a transformation across a vector field, which maps points near the center to the corners and points near the corners to the center, followed by a 3x3 blur. The blur causes each output pixel to depend on multiple input pixels, which allows for the possibility of work duplication.}
		\label{fig:flip-blur-1}
	\end{subfigure} &
	\begin{subfigure}[b]{0.45\textwidth}
		\begin{center}
		\includegraphics[width=0.85\textwidth]{canoe-3.png}
		\end{center}
		\caption{Final output. Stage 2 is altered to only apply the vector field to pixels when their colour value is darker than a certain threshold. This condition, which depends on a nondeterministic property of the input image, prevents Halide's optimizations from reasoning about the data dependencies between the two stages.}
		\label{fig:flip-blur-2}
	\end{subfigure}
	\end{tabular}
	\caption{Our sample image processing program consists of two stages, \texttt{invert()} (shown in \ref{fig:invert}) and \texttt{flip()} (shown in \ref{fig:flip-blur-1} and \ref{fig:flip-blur-2}). We designed this program expressly to create dynamic data dependencies with wide ranges across the image, to ensure a static function schedule could not avoid duplicating work.}

	\label{fig:flip}
\end{figure}
